# Introduction

Isoreader supports various dual inlet, continuous flow, and scan file formats. This vignette shows how to get started reading these raw IRMS data files and exporting the information to Excel. For more details on isoreader functionality and each file type, please read the **Full Examples** vignettes. For more information on downstream processing with isoverse, check out the [isoprocessor](https://isoprocessor.isoverse.org) package.

```{r, message=FALSE}
# load isoreader package
library(isoreader)
```

# Data files

For demonstration purposes, this vignette simply reads all supported dual inlet, continuous flow, and scan files that are bundled with the isoreader package.

```{r}
# all available examples
iso_get_reader_examples() %>% rmarkdown::paged_table()
```

# Dual Inlet Files

## Read

```{r}
# read all available examples
di_files <- iso_read_dual_inlet(iso_get_reader_examples_folder())
#di_files #shows all the files and their information
#iso_get_problems(di_files) #shows what is wrong with the files
```

## Save

```{r}
# save as r data storage (read back in with iso_read_dual_inlet)
iso_save(di_files, filepath = "data_output/di_save")
di_files <- iso_read_dual_inlet("data_output/di_save.di.rds")
#rds file = R Data Storage file. Good way to share files with others and store files
```

## Export

```{r}
# export to excel
iso_export_to_excel(di_files, filepath = "data_output/di_export")
```

# Continuous Flow Files

## Read

```{r}
# read all available examples
cf_files <- iso_read_continuous_flow(iso_get_reader_examples_folder())
```

## Save

```{r}
# save as r data storage (read back in with iso_read_continuous_flow)
iso_save(cf_files, filepath = "data_output/cf_save")
cf_files <- iso_read_continuous_flow("data_output/cf_save.cf.rds")
```

## Export

```{r}
# export to excel
iso_export_to_excel(cf_files, filepath = "data_output/cf_export")
```

# Scan Files

## Read

```{r}
# read all available examples
scan_files <- iso_read_scan(iso_get_reader_examples_folder())
```

## Save

```{r}
# save as r data storage (read back in with iso_read_scan)
iso_save(scan_files, filepath = "data_output/scan_save")
scan_files <- iso_read_scan("data_output/scan_save.scan.rds")
```

## Export

```{r}
# export to excel
iso_export_to_excel(scan_files, filepath = "data_output/scan_export")
```

# Visualization

```{r, message=FALSE}
library(isoprocessor)
library(tidyverse)
```

```{r}
# continuous flow files
cf_files
```

## Pick relevant files & information
##Filter out the files you are actually interested in

```{r}
cf_files %>% iso_get_file_info()
```


```{r}
my_cf_files <- cf_files %>% 
  # filter files
  iso_filter_files(str_detect(file_id, "continuous")) %>% 
  # set peak table
  iso_set_peak_table_from_auto_vendor_data_table() %>% 
  # select relevant file info
  iso_select_file_info(
    file_datetime, 
    analysis = Analysis, id1 = `Identifier 1`, id2 = `Identifier 2`, 
    prep = Preparation, comment = Comment
  ) %>% 
  # converting mV to V and mVs to Vs
  iso_convert_peak_table_units(V = mV, Vs = mVs)

my_cf_files
peak_tables <- my_cf_files %>% iso_get_peak_table() %>%  iso_make_units_explicit() #makes units obsvious within the table
view(peak_tables)
```

## Plot

```{r}
iso_plot_continuous_flow_data(my_cf_files)
```

## Styling the data
```{r}
?iso_plot_continuous_flow_data #gives you all the options for plotting continuous flow data
iso_plot_continuous_flow_data(
  iso_files = my_cf_files,
  data = c("2","3"), #selecting just for mass 2 and 3
  time_interval = c(500,800),
  color = data,
  panel = id1,
  peak_label = iso_format(rt, signif = 2),
  peak_label_size = 5,
  peak_label_filter = rt < 700,
  peak_marker = TRUE, #makes dashed line on middle of peak
  peak_bounds = TRUE #make gray bar across width of area integrated
  ) 

```




```{r}
iso_plot_continuous_flow_data(
  iso_files = my_cf_files,
  data = c("2", "3"),
  time_interval = c(700, 900),
  time_interval_units = "s",
  panel = NULL,
  color = data,
  peak_label = iso_format(rt),
  peak_label_size = 5,
  peak_bounds = TRUE,
  peak_marker = TRUE, 
  peak_label_filter = data == "2 [mV]",
)

ggsave("fig_output/my_fancy_chromatogram.pdf")
```

```{r, warning=FALSE, message=FALSE}
library(plotly)
ggplotly(dynamicTicks = TRUE)
```



